---
title: "Suggested Rates 8/22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
  
  
#Query lane bids joined to award info (only final round)
#This query joins bid data to award data, limited to only the data from the final round. Currently set to pull all lanes with a customer submit date since 6/10/2022. I then partition at 7/20 so there are 40 days of caster and 43 days of smasher data
```{r Andrew_query}
#I added super regions
#' postgres connect, pull, and disconnect from db
#'
#' @param tablename string tablename
#' @param dbname string dbname
#' @param user string username
#' @param pw string password
#' @param port num port
#' @param host string host
#' @param schema string schema if blank will use default schema
#' @param statement what do you want to read
#'
#' @return dataframe of table you wanted to pull
#' @export
#'
#' @examples
#' 

 
connect_pull_close <- function(tablename, 
                               dbname = "", 
                               user = "",
                               pw = "", 
                               port = , 
                               host = "",
                               schema = , 
                               statement = paste0("select * from ", tablename)) {
  if (!is.null(schema)) {
    options <- paste0("-c search_path=", schema)
  }
  conn <- tryCatch(
    {
      if (!is.null(schema)) {
        RPostgres::dbConnect(
          drv = RPostgres::Postgres(), 
          dbname = dbname, 
          host = host,
          port = port, 
          user = user, 
          password = pw, 
          options = options,
          sslmode = "require"
        )
      } else {
        RPostgres::dbConnect(
          drv = RPostgres::Postgres(), 
          dbname = dbname, 
          host = host,
          port = port, 
          user = user, 
          password = pw, 
          sslmode = "require"
        )
      }
    },
    error = function(cond) {
      message(cond)
      return(NA)
    }
  )
  ## get data from db
  data <- tryCatch(
    {
      RPostgres::dbGetQuery(conn = conn, statement = statement)
    },
    error = function(cond) {
      message(cond)
      return(NA)
    }
  )
  ## disconnect from db
  RPostgres::dbDisconnect(conn)
  return(data)
}

DB_PARAMS <- list(
    "dbname" = "contractual_prd",
    "user"  = "goldzoneingestion",
    "pw"  = "E2K8nz0KsDjyJcmjrejYVM7e",
    "host"   = "centralus-contractual-postgres-flex-db-prd-fpg.postgres.database.azure.com",
    "port"  = 5432,
    "schema" = "microfocals"
  )
statement <- "with lastround as (
select b.bid_id, max(b.round) as roundreporting
from microfocals.bids as b
--where b.expect_award='True'
group by b.bid_id
)
SELECT
b.bid_id
,b.id
,b.name
,b.round
,b.customer_submit_date as last_submit_date
,case when a.bid_id is null then 0 else 1 end as award_rec_flag
,cast(a.\"timestamp\" as date) as award_upload_date
,b.ear2
,b.ccode
,b.pricing_manager
,b.expect_award
,b.contract_id
,b.fratecaster_run_date
,c.bid_effective_date
,c.bid_expiration_date
,c.classification
,n.lane_id
,n.bid_lane_id
,n.origin_super_region
,n.destination_super_region
,n.origin_hierarchy
,n.origin_city
,n.origin_state
,concat(n.origin_city,', ',n.origin_state) as Origin
,n.origin_zip
,n.destination_hierarchy
,n.destination_city
,n.destination_state
,concat(n.destination_city,', ',n.destination_state) as Destination
,n.destination_zip
,n.miles
,n.mode
,n.bid_volume as available_volume
,case when n.submitted_allinrate >0 then n.bid_volume else 0 end as bid_volume
,case when n.submitted_allinrate >0 then n.bid_volume * n.submitted_allinrate else 0 end as gross_revenue_bid
,case when n.submitted_allinrate >0 then n.bid_volume * (n.submitted_allinrate - ((coalesce(n.forecasted_coh,n.fratecaster_50th_cost)+n.fuel_surcharge)*n.miles)) else 0 end as agp_bid
,n.submitted_allinrate
,n.\"submitted_RPM\"
,n.starting_rate_guidance
,n.lane_strategy
--,n.coh_percentile
,n.forecasted_coh
,n.fratecaster_15th_cost
,n.fratecaster_50th_cost
,n.fratecaster_85th_cost
,case when al.awarded_volume>0 then 1 else 0 end as awarded_lane
,al.awarded_rpm
,al.awarded_volume
,n.\"chain_3_month_FmicLineHaulCost\" as chain_3month
,n.is_drop_trailer
,n.is_hazmat
,n.is_high_value
from lastround l
join microfocals.bids as b on l.bid_id=b.bid_id and l.roundreporting=b.round
left join microfocals.contracts as c on b.contract_id = c.contract_id
join microfocals.lanes as n on b.bid_id = n.bid_id and l.roundreporting=n.\"round\"
left join microfocals.awards as a on a.bid_id = b.bid_id
left join microfocals.awarded_lanes as al on a.id = al.award_id_fk and n.bid_lane_id = al.bid_lane_id
where b.customer_submit_date>='2022-06-15 00:00:00'
"

df <- connect_pull_close(
    dbname = DB_PARAMS[["dbname"]], 
    user = DB_PARAMS[["user"]],
    pw = DB_PARAMS[["pw"]], 
    port = DB_PARAMS[["port"]], 
    host = DB_PARAMS[["host"]],
    schema = "load_data", 
    statement = statement
  )

statement <- "select * from table"
```

#Filter out incomplete data and add boolean variables
```{r df_added_cases}
#1. Removed cases without a starting_rate_guidance or submitted_RPM, since the goal is to compare starting_rate_guidance to submitted_RPM and how that affects what we win
#2. Created a series of booleans to make it easy to filter.
# -submit_is_award: is the submitted rate equal to the awarded rate in a lane tagged as awarded?
# -suggest_is_award: is the suggested rate (starting_rate_guidance) equal to the awarded rate in a lane tagged as awarded?
# -suggest_is_submit: is the suggested rate (starting_rate_guidance) equal to the submitted rate?
# -suggest_versus_submit: How does the suggested rate compare to the rate submitted by the TPMs?
# -suggest_versus_award: How does the suggested rate compare to the rate awarded?
#3. Converted categorical variables to factors

df_joined <- df %>% 
  filter(!is.na(starting_rate_guidance),
         !is.na(submitted_RPM),
         round == 1) %>%
  mutate(
    submit_is_award = case_when(
      submitted_RPM == awarded_rpm & awarded_lane == 1 ~ TRUE,
      submitted_RPM != awarded_rpm | awarded_lane == 0 | is.na(awarded_rpm) ~ FALSE
      ),
    suggest_is_award = case_when(
      starting_rate_guidance == awarded_rpm & awarded_lane == 1 ~ TRUE,
      starting_rate_guidance != awarded_rpm | awarded_lane == 0| is.na(awarded_rpm) ~ FALSE
    ),
    suggest_versus_submit = case_when(
      starting_rate_guidance > submitted_RPM ~ 'Higher',
      starting_rate_guidance == submitted_RPM ~ 'Equal',
      starting_rate_guidance < submitted_RPM ~ 'Lower'
    ),
    suggest_versus_award = case_when(
      starting_rate_guidance > awarded_rpm ~ 'Higher',
      starting_rate_guidance == awarded_rpm ~ 'Equal',
      starting_rate_guidance < awarded_rpm ~ 'Lower'
    ),
    round = as_factor(round),
    award_rec_flag = as_factor(award_rec_flag),
    expect_award = as_factor(expect_award),
    classification = as_factor(classification),
    origin_state = as_factor(origin_state),
    destination_state = as_factor(destination_state),
    lane_strategy = as_factor(lane_strategy),
    is_drop_trailer = as_factor(is_drop_trailer),
    is_hazmat = as_factor(is_hazmat),
    is_high_value = as_factor(is_high_value),
    suggest_versus_award = as_factor(suggest_versus_award),
    suggest_versus_submit = as_factor(suggest_versus_submit),
    mode = as.factor(mode)
  )
df_joined <- df_joined %>%
  mutate(smasher = case_when(
    fratecaster_run_date < '2022-07-20' ~ "fRateCaster",
    fratecaster_run_date >= '2022-07-20' ~ "fRateSmasher")) %>%
  filter(!mode %in% c("F","Flatbed"),
         is_drop_trailer == FALSE,
         origin_super_region != "Canada",
         destination_super_region != "Canada")
```

# Summary (as of 8/19): 6-1-2022 forward
- 189354 [total_lanes] round 1 lanes with 565 [unique_bid_id] unique bid_id
- all have starting_rate_guidance and submitted_RPM != 0
- Of all these lanes, we were awarded 12.2% [award_percentage] of them
- Of all these lanes, the submitted rate was the awarded rate 11.22% [submit_award_rate] of the time
- Of all these lanes, the starting rate guidance was the awarded rate 5.8% [suggest_award_rate] of the time

- Of the lanes we were awarded (that 12.2%):
  - The starting_rate_guidance was the award 47.8% [suggest_award_won] of the time
  
```{r basic_comparisons}
total_lanes <- nrow(df_joined)
unique_bid_id <- length(unique(df_joined$bid_id))
award_percentage <- mean(df_joined$awarded_lane)
submit_award_rate <- mean(df_joined$submit_is_award)
suggest_award_rate <- mean(df_joined$suggest_is_award)

awarded_lanes <- df_joined %>% filter(awarded_lane == 1)
suggest_award_won <- mean(awarded_lanes$suggest_is_award)
```

108966 lanes over 323 bid id run before 7-20
- 10.17% win rate
- the rate submitted was the awarded rate 8.54% of the time
- the rate suggested was the awarded rate 5.85% of the time
- of the lanes we were awarded, the suggested rate was the award 57.5% of the time


89280 lanes over 277 run on or after 7-20
- 9.83% win rate
- the rate submitted was the awarded rate 9.3% of the time
- the rate suggested was the awarded rate 3.17% of the time
- of the lanes we were awarded, the suggested rate was the award 32.26% of the time

```{r compare pre_post_smasher}
cast_lanes <- df_joined %>% filter(smasher == "fRateCaster")
nrow(cast_lanes)
length(unique(cast_lanes$bid_id))
mean(cast_lanes$awarded_lane)
mean(cast_lanes$submit_is_award)
mean(cast_lanes$suggest_is_award)
awarded__cast_lanes <- cast_lanes %>% filter(awarded_lane == 1)
mean(awarded__cast_lanes$suggest_is_award)


smash_lanes <- df_joined %>% filter(smasher == "fRateSmasher")
nrow(smash_lanes)
length(unique(smash_lanes$bid_id))
mean(smash_lanes$awarded_lane)
mean(smash_lanes$submit_is_award)
mean(smash_lanes$suggest_is_award)
awarded__smash_lanes <- smash_lanes %>% filter(awarded_lane == 1)
mean(awarded__smash_lanes$suggest_is_award)
```

```{r suggested_becomes_award_BASIC}
#Goal: Visualize Percentages described above
#Lanes where suggest was award or not
ggplot(df_joined %>% drop_na(suggest_is_award)) +
  geom_bar(aes(suggest_is_award)) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Suggested Rate == Awarded Rate", y = "Lanes", title = "Occurences of Suggested RPM Becoming Awarded RPM")

#lanes where suggest was award or not split by smasher
ggplot(df_joined %>% drop_na(suggest_is_award)) +
  geom_bar(aes(suggest_is_award)) +
  facet_wrap(~smasher) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Suggested Rate == Awarded Rate", y = "Lanes", title = "Occurences of Suggested RPM Becoming Awarded RPM")

#lanes where suggest was award, just of lanes we won
ggplot(df_joined %>% drop_na(suggest_is_award) %>% filter(awarded_lane==1))+
  geom_bar(aes(suggest_is_award)) +
  facet_wrap(~smasher) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Suggested Rate == Awarded Rate", y = "Awarded Lanes", title = "Occurences of Suggested RPM Becoming Awarded RPM", subtitle = "Of Awarded Lanes")

ggplot(df_joined %>% drop_na(suggest_is_award) %>% filter(awarded_lane==1))+
  geom_bar(aes(suggest_is_award, fill = smasher), position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Suggested Rate == Awarded Rate", y = "Awarded Lanes", title = "Occurences of Suggested RPM Becoming Awarded RPM", subtitle = "Of Awarded Lanes", fill = "Model") +
  theme_minimal()


ggplot(df_joined %>% drop_na(suggest_is_award) %>% filter(awarded_lane==1) %>% group_by(smasher) %>% summarise(count = sum(awarded_lane), prop = mean(suggest_is_award)))+
  geom_col(aes(x = smasher, y = count, fill = prop), position = "dodge") +
  geom_text(aes(x = smasher, y = count - 1000, label = scales::percent(round(prop, 4))), color = "white", size = 5) +
  scale_y_continuous(labels = scales::comma)  +
  scale_fill_continuous(low = "dark red", high = "dark green") +
  theme_minimal() +
  labs(x = "Model", y = "Lane Count", fill = "Awarded Rate\nis Suggested\nRate", title = "Proportion of Awarded Lanes that Won with Suggested Rate")
```


#Look at where suggested rates of single-shot data are used by mode, super region
```{r  mode}
#Ok,  let's look at how we are awarded lanes by mode. Then we can see if suggested behavior is just following awarded behavior:
ggplot(df_joined %>% drop_na(awarded_lane)) +
  geom_bar(aes(as.factor(awarded_lane), fill = smasher), position = "dodge") +
  facet_wrap(~mode, scales = "free_y", labeller = labeller(mode = c("F" = "Flatbed", "R" = "Reefer", "V" = "Van")))+
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Awarded Lane", y = "Lanes", title = "Lane Outcome by Mode")+
  theme(legend.title = element_blank())  
#fRatesmasher is winning more reefer and less van than caster

#Let's look at the proportion of those awarded lanes by mode. i.e. what is our awarded lane rate by group?
prop_won_by_mode <- df_joined %>% group_by(mode) %>% summarise(prop = mean(awarded_lane))

ggplot(prop_won_by_mode) +
  geom_col(aes(x= as.factor(mode), y =prop), position = "dodge") +
  geom_text(aes(x = as.factor(mode), y = prop - 0.003, label = scales::percent(prop)), color = "white", size =5) +
  labs(x = "Mode", y= "Awarded Lanes", title = "Awarded Lanes by Mode") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  scale_x_discrete(labels = c("Reefer", "Van"))+
  theme(legend.title = element_blank())

#Goal: The above plot was more challenging to look at based on count disparities, let's look at percentages where suggested rpm was the awarded rpm split by mode:
prop_suggest_is_award_mode <- df_joined %>% group_by(mode) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= mode, y =prop)) +
  geom_text(aes(x = mode, y=prop + 0.0025, label = scales::percent(round(prop,4))), color = "black", size = 5)+
  labs(x = "Mode", y= "Percentage of Lanes", title = "Suggested_RPM becomes Awarded_RPM") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  scale_x_discrete(labels = c( "Reefer", "Van"))+
  theme(legend.title = element_blank())


#Goal: Let's see if the behavior by mode varies between models

prop_suggest_is_award_mode <- df_joined %>% group_by(mode, smasher) %>% summarise(prop = mean(awarded_lane, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= mode, y =prop, fill = smasher), position = "dodge") +
  labs(x = "Mode", y= "Percentage of Lanes", title = "Award Rate") +
  geom_text(aes(x= c(0.8, 1.2, 1.75, 2.2), y =prop + 0.0025, label = scales::percent(round(prop,4))) )+
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  scale_x_discrete(labels = c( "Reefer", "Van"))+
  theme(legend.title = element_blank())

prop_suggest_is_award_mode <- df_joined %>% filter(awarded_lane == 1) %>% group_by(mode) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= mode, y =prop), position = "dodge") +
  labs(x = "Mode", y= "Percentage of Lanes", title = "Percentage of Awarded Lanes where Suggested_RPM is Awarded_RPM") +
  geom_text(aes(x= mode, y =prop - 0.025, label = scales::percent(round(prop,4))), color = "white", size =5 )+
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  scale_x_discrete(labels = c( "Reefer", "Van"))+
  theme(legend.title = element_blank())

prop_suggest_is_award_mode <- df_joined %>% filter(awarded_lane == 1) %>% group_by(mode, smasher) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= mode, y =prop, fill = smasher), position = "dodge") +
  labs(x = "Mode", y= "Percentage of Lanes", title = "Percentage of Awarded Lanes where Suggested_RPM is Awarded_RPM") +
  geom_text(aes(x= c(0.8, 1.2, 1.78, 2.23), y =prop - 0.025, label = scales::percent(round(prop,4))) )+
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  scale_x_discrete(labels = c( "Reefer", "Van"))+
  theme(legend.title = element_blank())

```

```{r origin_super_region}
#Ok,  let's look at how we are awarded lanes across regions:
ggplot(df_joined %>% drop_na(awarded_lane)) +
  geom_bar(aes(as.factor(awarded_lane), fill = smasher), position = "dodge") +
  facet_wrap(~origin_super_region, scales = "free_y")+
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Awarded Lane", y = "Lanes", title = "Lane Outcome by Origin Super Region")+
  theme(legend.title = element_blank())  

#Let's look at the proportion of those awarded lanes. i.e. what is our awarded lane rate by group?
prop_won_by_mode <- df_joined %>% group_by(origin_super_region) %>% summarise(prop = mean(awarded_lane))

ggplot(prop_won_by_mode) +
  geom_col(aes(x= reorder(origin_super_region,prop), y =prop), position = "dodge") +
  geom_text(aes(x = reorder(origin_super_region,prop), y = prop - 0.003, label = scales::percent(prop)), color = "white", size =5) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(x = "Origin Super Region", y = "Award Rate", title = "Percentage of Lanes Awarded by Origin Super Region")

#How are those proportions different between models?
prop_won_by_mode <- df_joined %>% group_by(origin_super_region, smasher) %>% summarise(prop = mean(awarded_lane))

ggplot(prop_won_by_mode) +
  geom_col(aes(x= reorder(origin_super_region,prop), y =prop, fill = smasher), position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(x = "Origin Super Region", y = "Award Rate", title = "Percentage of Lanes Awarded by Origin Super Region")


#Goal: Look at percentages where suggested rpm was the awarded rpm split by osr:
prop_suggest_is_award_mode <- df_joined %>% filter(awarded_lane == 1) %>% group_by(origin_super_region) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= reorder(origin_super_region,prop), y =prop)) +
  geom_text(aes(x = reorder(origin_super_region,prop), y=prop - 0.01, label = scales::percent(round(prop,4))), color = "white", size = 5)+
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(x = "Origin Super Region",y= "Percentage of Lanes", title = "Percentage of Awarded Lanes where the Suggested Rate Wins")

#Let's divide by model
prop_suggest_is_award_mode <- df_joined %>% filter(awarded_lane == 1) %>% group_by(origin_super_region, smasher) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= reorder(origin_super_region,prop), y =prop, fill = smasher), position="Dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(x = "Origin Super Region",y= "Percentage of Lanes", title = "Percentage of Awarded Lanes where the Suggested Rate Wins")

```

```{r destination_super_region}
#Ok,  let's look at how we are awarded lanes across regions:
ggplot(df_joined %>% drop_na(awarded_lane)) +
  geom_bar(aes(as.factor(awarded_lane), fill = smasher), position = "dodge") +
  facet_wrap(~destination_super_region, scales = "free_y")+
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Awarded Lane", y = "Lanes", title = "Lane Outcome by Destination Super Region")+
  theme(legend.title = element_blank())  

#Let's look at the proportion of those awarded lanes. i.e. what is our awarded lane rate by group?
prop_won_by_mode <- df_joined %>% group_by(destination_super_region) %>% summarise(prop = mean(awarded_lane))

ggplot(prop_won_by_mode) +
  geom_col(aes(x= reorder(destination_super_region,prop), y =prop), position = "dodge") +
  geom_text(aes(x = reorder(destination_super_region,prop), y = prop - 0.003, label = scales::percent(prop)), color = "white", size =5) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(x = "Destination Super Region", y = "Award Rate", title = "Percentage of Lanes Awarded by Destination Super Region")

#How are those proportions different between models?
prop_won_by_mode <- df_joined %>% group_by(destination_super_region, smasher) %>% summarise(prop = mean(awarded_lane))

ggplot(prop_won_by_mode) +
  geom_col(aes(x= reorder(destination_super_region,prop), y =prop, fill = smasher), position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(x = "Destination Super Region", y = "Award Rate", title = "Percentage of Lanes Awarded by Destination Super Region")


#Goal: Look at percentages where suggested rpm was the awarded rpm split by osr:
prop_suggest_is_award_mode <- df_joined %>% filter(awarded_lane == 1) %>% group_by(destination_super_region) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= reorder(destination_super_region,prop), y =prop)) +
  geom_text(aes(x = reorder(destination_super_region,prop), y=prop - 0.01, label = scales::percent(round(prop,4))), color = "white", size = 5)+
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(x = "Destination Super Region",y= "Percentage of Lanes", title = "Percentage of Awarded Lanes where the Suggested Rate Wins")

#Let's divide by model
prop_suggest_is_award_mode <- df_joined %>% filter(awarded_lane == 1) %>% group_by(destination_super_region, smasher) %>% summarise(prop = mean(suggest_is_award, na.rm=TRUE))

ggplot(prop_suggest_is_award_mode) +
  geom_col(aes(x= reorder(destination_super_region,prop), y =prop, fill = smasher), position="Dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(x = "Destination Super Region",y= "Percentage of Lanes", title = "Percentage of Awarded Lanes where the Suggested Rate Wins") +
  geom_hline(aes(yintercept =overall[1,] ))

overall <- df_joined %>% filter(awarded_lane == 1) %>%summarise(prop = mean(suggest_is_award, na.rm=TRUE))
```



#Start of suggested rates & geography overview
```{r create_list_of_us_states}
#I want to remove non-US states
list_of_us <- unique(df_joined$origin_state) 
list_of_us <- list_of_us[-c(21, 44, 45, 48,54, 56, 57, 58, 59, 60, 61, 62, 63, 64)]

#Calculate the percentage of awarded lanes that won using suggested rate across all awarded lanes
stats <- df_joined %>% filter(awarded_lane == 1) %>% summarise(avg = mean(suggest_is_award))
avg <- stats[,1]
```

```{r origin_state_suggest_is_award}
#Out of awarded lanes originating in the US, what percentage of awards used the suggested rate, partioned by origin state

percent_suggest_of_award <- df_joined %>% filter(awarded_lane == 1, ) %>% group_by(origin_state) %>% summarise(prop_of_award = mean(suggest_is_award), sum = n()) %>% arrange(sum)

#Visualize it with a line showing the overall rate at which awarded lanes use the suggested rate
ggplot(percent_suggest_of_award %>% filter(sum > 30) ) +
  geom_col(aes(x= reorder(origin_state,prop_of_award), y = prop_of_award))+
  scale_y_continuous(labels = scales::percent) +
  labs(x = "origin_state", y = "Percentage of Awarded Lanes",  title = "Percentage of Awarded Lanes That Won With Suggested Rate")+
  theme_minimal() +
  geom_abline(slope = 0, intercept = avg, color = "red")

#it looks like mayber there are some extremes in the orgin_state breakdown
```



```{r destination_state_suggest_is_award}
#awarded lanes with suggested rate partioned by destination state
#Exact same analysis as chunk above, just with destination state
percent_suggest_of_award <- df_joined %>% filter(awarded_lane == 1, destination_state %in% list_of_us) %>% group_by(destination_state) %>% summarise(prop_of_award = mean(suggest_is_award), sum = length(destination_state)) %>% arrange(sum)

ggplot(percent_suggest_of_award %>% filter(sum > 30)) +
  geom_col(aes(x= reorder(destination_state,prop_of_award), y = prop_of_award))+
  scale_y_continuous(labels = scales::percent) +
  labs(x = "destination_state", y = "Percentage of Awarded Lanes",  title = "Percentage of Awarded Lanes That Won With Suggested Rate")+
  theme_minimal()  +
  geom_abline(slope = 0, intercept = avg, color = "red")
```

# I'm going to switch gears.
#Which bid-level or lane-level features are different between awarded lanes using suggested rate and awarded lanes that didn't use suggested rate?

```{r}
awarded_lanes <- df_joined %>% filter(awarded_lane == 1) 
#What if we do a linear regression on whether or not the suggested rate was the award?

suggest_awarded_glm <- glm(suggest_is_award ~  
                             classification + 
                             #origin_state + 
                             #destination_state + 
                             #miles + 
                             mode + 
                             #bid_volume + 
                             #submitted_allinrate + 
                             #submitted_RPM +            
                             #starting_rate_guidance + 
                             lane_strategy +
                             forecasted_coh +
                             #fratecaster_50th_cost + 
                             is_drop_trailer ,
                             #is_hazmat + 
                             #suggest_versus_submit , 
                           data = awarded_lanes, 
                           family = "binomial")

summary(suggest_awarded_glm)
regclass::VIF(suggest_awarded_glm)
```
#takeaways:
- RePrice less likely suggest is awarded, Minibid more liekly
- Reefer, Van, more likely to take suggested
- increase in submitted_rpm decreases liklihood suggest is taken
- backup and route guide increase
- forecasted coh increases
- drop trailer increases

```{r mode}
awarded_lanes <- awarded_lanes %>% mutate(mode = as.factor(mode))
ggplot(awarded_lanes %>% drop_na(suggest_versus_award))+
  geom_bar(aes(mode, fill=suggest_versus_award), position = "dodge") +
    theme_minimal() +
  labs(x = "Mode", y="Lane Count", fill = "Suggested Rate\nvs Award", title = "Awarded Lanes By Mode")

#Lets look at it by proportion
granular <- awarded_lanes %>% 
  group_by(mode, suggest_versus_award) %>% 
  summarise(num = n(),
            total = nrow(mode))

groups <- awarded_lanes %>% 
  group_by(mode) %>% 
  summarise(num = n())

get_prop <- left_join(granular, groups, by = "mode") %>% mutate(prop = num.x/num.y)
  

ggplot(get_prop %>% drop_na(suggest_versus_award)) +
  geom_col(aes(x=mode, y=prop, fill = suggest_versus_award), position = "dodge")+
  scale_y_continuous(label=scales::percent) +
  labs(x = "Mode",y="Percentage of Awarded Lanes", fill = "Suggested Vs.\nAwarded Rate", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Split by Mode") +
  scale_x_discrete(labels = c("Flatbed", "Reefer", "Van")) +
  theme_minimal()
```

```{r classification}

ggplot(awarded_lanes %>% filter(!classification %in% c("Project", "NonAwardBid", "Outsource")) %>% drop_na(suggest_versus_award))+
  geom_bar(aes(classification, fill=suggest_versus_award), position = "dodge") +
    theme_minimal() +
  labs(x = "Classification", y="Lane Count", fill = "Suggested Rate\nvs Award", title = "Awarded Lanes By Classification")


#lets check the above by percentages

awarded_lanes %>% group_by(classification) %>% filter(!classification %in% c("Project", "NonAwardBid", "Outsource")) %>% summarise(prop_suggest_awarded = mean(suggest_is_award, na.rm=TRUE)) %>% ggplot() + geom_col(aes(x=classification, y=prop_suggest_awarded)) +
  scale_y_continuous(label = scales::percent) +
  labs(x = "Classification", y = "Percentage of Awarded Lanes", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Grouped by Classification") +
  theme_minimal()

#Lets look at it by proportion
granular <- awarded_lanes %>% filter(!classification %in% c("Project", "NonAwardBid", "Outsource")) %>%
  group_by(classification, suggest_versus_award) %>% 
  summarise(num = n(),
            total = nrow(mode))

groups <- awarded_lanes %>% filter(!classification %in% c("Project", "NonAwardBid", "Outsource")) %>%
  group_by(classification) %>% 
  summarise(num = n())

get_prop <- left_join(granular, groups, by = "classification") %>% mutate(prop = num.x/num.y)
  

ggplot(get_prop %>% drop_na(suggest_versus_award)) +
  geom_col(aes(x=classification, y=prop, fill = suggest_versus_award), position = "dodge")+
  scale_y_continuous(label=scales::percent) +
  labs(x = "Classification",y="Percentage of Awarded Lanes", fill = "Suggested Vs.\nAwarded Rate", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Split by Classification") +
  theme_minimal()
```

```{r lane_strategy}
awarded_lanes <- awarded_lanes %>% 
  mutate(lane_strategy = fct_collapse(lane_strategy,
    Targeted = c("Targeted", "Target"),
    Backup = c("Backup", "backup"),
    `Route Guide` = c("Route Guide")))

ggplot(awarded_lanes %>% drop_na(suggest_versus_award)) +
  geom_bar(aes(lane_strategy, fill = suggest_versus_award), position = "dodge") +
  theme_minimal() +
  labs(x = "Lane Strategy", y = "Lane Count", fill = "Suggested Rate\nvs Award", title = "Awarded Lanes by Lane Strategy")

awarded_lanes %>% group_by(lane_strategy) %>%  summarise(prop = mean(suggest_is_award)) %>% ggplot()+geom_col(aes(x= lane_strategy, y=prop)) +
  theme_minimal() +
  labs(x = "Lane Strategy", y="Percentage of Awarded Lanes", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Grouped by Lane Strategy")

#Lets look at it by proportion
granular <- awarded_lanes %>%
  group_by(lane_strategy, suggest_versus_award) %>% 
  summarise(num = n(),
            total = nrow(lane_strategy))

groups <- awarded_lanes %>% 
  group_by(lane_strategy) %>% 
  summarise(num = n())

get_prop <- left_join(granular, groups, by = "lane_strategy") %>% mutate(prop = num.x/num.y)
  

ggplot(get_prop %>% drop_na(suggest_versus_award)) +
  geom_col(aes(x=lane_strategy, y=prop, fill = suggest_versus_award), position = "dodge")+
  scale_y_continuous(label=scales::percent) +
  labs(x = "Lane Strategy",y="Percentage of Awarded Lanes", fill = "Suggested Vs.\nAwarded Rate", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Split by Lane Strategy") +
  theme_minimal()
```

#lets ignore drop trailer
```{r drop_trailer}
ggplot(awarded_lanes) +
  geom_bar(aes(is_drop_trailer, fill = suggest_versus_award), position = "dodge") +
  theme_minimal() +
  labs(x = "Drop Trailer", y = "Lane Count", fill = "Suggested Rate\nAwarded", title = "Awarded Lanes by Lane Strategy")

awarded_lanes %>% group_by(is_drop_trailer) %>%  summarise(prop = mean(suggest_is_award)) %>% ggplot()+geom_col(aes(x= is_drop_trailer, y=prop)) +
  theme_minimal() +
  labs(x = "Is Drop Trailer", y="Percentage of Awarded Lanes", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Grouped by is_drop_trailer") +
  scale_y_continuous(label = scales::percent)

#Lets look at it by proportion
granular <- awarded_lanes %>%
  group_by(is_drop_trailer, suggest_versus_award) %>% 
  summarise(num = n(),
            total = nrow(is_drop_trailer))

groups <- awarded_lanes %>% 
  group_by(is_drop_trailer) %>% 
  summarise(num = n())

get_prop <- left_join(granular, groups, by = "is_drop_trailer") %>% mutate(prop = num.x/num.y)
  

ggplot(get_prop %>% drop_na(suggest_versus_award)) +
  geom_col(aes(x=is_drop_trailer, y=prop, fill = suggest_versus_award), position = "dodge")+
  scale_y_continuous(label=scales::percent) +
  labs(x = "Drop Trailer",y="Percentage of Awarded Lanes", fill = "Suggested Vs.\nAwarded Rate", title = "Percentage of Lanes Awarded with Suggested Rate", subtitle = "Split by Drop Trailer") +
  theme_minimal()
``` 

